---
title: "Manual of fastANCOM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{manual}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

### Install and load fastANCOM

```{r}
if(!require(fastANCOM)) {
  devtools::install_github('ZRChao/fastANCOM')  
}
library(fastANCOM)
packageVersion('fastANCOM')
browseVignettes('fastANCOM')
```


### Example

We take the differential abundance analysis of 2 groups for illustration.
The data was generated from Poisson distribution for each group with 30 individuals and 100 microbes. We set the first 10 microbes to be differetial by multipling a effect size vector for the abundance of one group, which is the log2FC from a norm distribution $N(4, 1)$.

#### Data generation

```{r}
set.seed(123)
data <- matrix(rpois(100*60, 10), 60)
b <- rnorm(10, 4)
data[31:60,1:10] <- t(t(data[31:60,1:10])*2^b)
group <- rep(c(0, 1), each=30)
rownames(data) <- paste0('sample', 1:60)
colnames(data) <- paste0('microbe', 1:100)
names(b) <- paste0('microbe', 1:10)

# take a look of the log2FC
kable_classic(kbl(t(round(b,4)),align= "l",caption='log2FC for the first 10 microbes'),
              html_font = "Cambria")  

# take a look of the data
kable_classic(kbl(data[1:10, 1:6], align = "l",  caption='Data at a glance'), 
               full_width = T, html_font = "Cambria") 
```

#### Model fitting

Here other parameters are set as default.

```{r}
fit <- fastANCOM(Y=data, x=group)
kable_classic(kbl(summary(fit), caption = 'Results summary'), 
              full_width = T, html_font = "Cambria")
kable_classic(kbl(head(fit, 10), align = "l", 
        caption='Results of fastANCOM'),  full_width = T, html_font = "Cambria")
```


> __Take a look of the computation time with 100 replicates, fastANCOM thousands of times faster than original ANCOM__

```{r, warning=F, message=F}
system.time(replicate(100, fit <- fastANCOM(Y=data, x=group, effect = T)))
system.time(replicate(100, fit <- fastANCOM(Y=data, x=group, effect = F)))
system.time(replicate(1, fit <- ANCOM(Y=data, x=group, tfun = wilcox.test))) 
system.time(replicate(1, fit <- ANCOM(Y=data, x=group, tfun = t.test))) 
system.time(replicate(100, fit <- ANCOM(Y=data, x=group, tfun = 't2'))) 
```

#### Results with plot 

We present plots of reject number of each microbe, valcano plot, and the log2 fold changes estimation compared to the true settings.

```{r, fig.align='center', fig.width=6.5, fig.height=3.6}
library(ggplot2)
library(ggpubr)
wp <- ggplot(data=fit, aes(x=1:100,y=Reject.number, color=REJECT, shape=REJECT)) + 
  geom_bar(aes(y=Reject.number), stat = "identity", width = 0.005, color='lightgrey') +
  # geom_vline(xintercept = 1:100, linetype='dashed', size=0.1) +
  geom_point(size=2) + theme_linedraw() + 
  labs(x='Microbe',y='W', title='Reject number') +
  scale_colour_brewer(palette = "Set1", direction=-1) + 
  theme(legend.position = c(0.85, 0.78), 
        legend.background = element_rect(fill = "white", colour = "black"))
wp

vp <-ggplot(data=fit,aes(x=log2FC,y=-log10(log2FC.pval), color=REJECT, shape=REJECT)) +
  geom_hline(yintercept=-log10(0.05/nrow(fit)),
             linetype="dashed",color="black",size=0.5)+ 
  geom_point(size=2) + theme_linedraw() + labs(y='log2(p-value)', title='Volcano plot') +
  theme(legend.position='none') +
  scale_colour_brewer(palette = "Set1",direction=-1) 
vp

tmpb <- data.frame(Microbe=factor(1:10),log2FC=c(b, fit[fit$REJECT, 1]), 
                   se=c(rep(0, 10), fit[fit$REJECT, 2]),
                   Type=c(rep('Truth', 10), rep('Estimation', 10)))
ep <- ggplot(data=tmpb, aes(x=Microbe, y=log2FC, color=Type, shape=Type))  +
  geom_point(size=2, position =  position_dodge(0.2) ) + 
  geom_errorbar(aes(ymin=log2FC-2*se, ymax=log2FC+2*se),
                width=0.2, position =  position_dodge(0.2)) +
  theme_linedraw() + labs(title='Log2FC estimation') +
  scale_colour_brewer(palette = "Set1", direction=-1) +
   theme(legend.position = c(0.7, 0.8), 
        legend.background = element_rect(fill = "white", colour = "black"))
ep
```



```{r, fig.align='center', fig.width=8, fig.height=5, eval=F, echo=F}
figs <- ggarrange(wp, ggarrange(vp, ep, ncol=2), nrow=2)
figs
```


```{r}
sessionInfo()
```



